---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "Yuta Tsukumo"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
library(patchwork)

load("CRDT Data.RData")
ls()
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Cases)) + geom_point()
```



If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. Write code for a plot and describe what you observe?

```{r}
data <- covid_data_count
summary(data)
table(data$state)
# including PR;Puerto Rico, s in total 52 states are shown

# categorize race as (White, Black, Latinx, Asian, Other)
data$race_cat <- ifelse(data$race %in% c("Total", "White", "Black", "Latinx", "Asian"), data$race, "Other")
table(data$race_cat)
  
# categorize states into 4 regions (Northeast, Midwest, South, West)
# Puerto Rico included into South region
data <- data %>%
  mutate(
    region = ifelse(state_name %in% c(
      "Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island",
      "Vermont", "New Jersey", "New York", "Pennsylvania"
    ), "Northeast",
    ifelse(state_name %in% c(
      "Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin",
      "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota",
      "South Dakota"
    ), "Midwest",
    ifelse(state_name %in% c(
      "Delaware","District of Columbia", "Florida", "Georgia", "Maryland", "North Carolina",
      "South Carolina", "Virginia", "West Virginia", "Alabama", "Kentucky",
      "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas"
    ), "South",
    ifelse(state_name %in% c(
      "Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico",
      "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon",
      "Washington", "Puerto Rico"
    ), "West", "Unknown"
    ))))
  )

table(data$region)

# create short data set where case and deaath are summed up by region/race_cat
# How do I create a dataset with a date column, a race column, a region column, a number of infections column, and a number of deaths column, and then merge the columns so that for each date, if the race and region are the same, the number of infections and deaths are added together?


p_case <- data %>%
  filter(race_cat %in% c ("White", "Black", "Latinx", "Asian","Other")) %>%
  ggplot(aes(x = date, y = Cases, color = race_cat)) + 
  geom_line() +
  labs(title = "Cases_by_Race") +
  facet_wrap(~region) +
  theme_minimal()

p_case

p_death <- data %>%
  filter(race_cat %in% c ("White", "Black", "Latinx", "Asian","Other")) %>%
  ggplot(aes(x = date, y = Deaths, color = race_cat)) + 
  geom_line() +
  labs(title = "Cases_by_Race") +
  facet_wrap(~region) +
  theme_minimal()
p_death

```

2. Are there any conclusions you can draw?

In all four regions, both the number of cases and deaths are increasing from April 2020 to March 2021, with particularly sharp increases around summer 2020 and January 2021.
In all four regions, whites account for a larger proportion of cases and deaths when compared among races, but characteristically, in the West region, Latinos account for a larger proportion of cases than whites, and Latinos account for a larger proportion of deaths than in the other regions.

3. What were you not able to do due to not having the R knowledge?

The analysis after grouping by race and state does not proceed due to lacking my skill to create a data set that merges the rows so that the number of cases and deaths are added together if the race and region are the same on each date.

4. What other data would be useful to better understand this data?

With basic information such as age and gender, the comparison of infection trends in the patient background can be analyzed in more detail.
In addition, when considering the burden on medical institutions, we should collect not only the number of infections or deaths, but also the number of severely ill patients, which is important when considering policies for the maintenance and support of medical institutions.
Information on vaccination history would also be helpful.
